Lesson 5
========================================================

### Load data

```{r}
setwd('~/Git/data-analyst-nanodegree/P4/eda-course-materials/lesson5/') # ABSOLUTE PATH! CHANGE if PARENT directory changes!
getwd()

list.files()
pf <- read.csv('../lesson3/pseudo_facebook.tsv',sep = '\t')


```

### Multivariate Data
Notes:

***

### Moira Perceived Audience Size Colored by Age
Notes:

***

### Third Qualitative Variable
Notes:

```{r Third Qualitative Variable}
library(ggplot2)
 #ggplot(aes(x = gender, y = age),data = subset(pf, !is.na(gender))) + geom_histogram()
```
```{r}
# Write code to create a new data frame,
# called 'pf.fc_by_age_gender', that contains
# information on each age AND gender group.

# The data frame should contain the following variables:

#    mean_friend_count,
#    median_friend_count,
#    n (the number of users in each age and gender grouping)

# Here is an example of the structure of your data frame. Your
# data values will be different. Note that if you are grouping by
# more than one variable, you will probably need to call the
# ungroup() function. 

#   age gender mean_friend_count median_friend_count    n
# 1  13 female          247.2953                 150  207
# 2  13   male          184.2342                  61  265
# 3  14 female          329.1938                 245  834
# 4  14   male          157.1204                  88 1201

# See the Instructor Note for two hints.

# DO NOT DELETE THESE NEXT TWO LINES OF CODE
# ==============================================================
suppressMessages(library(dplyr))

# ENTER YOUR CODE BELOW THIS LINE.
# ==============================================================
library(dplyr)


pf.fc_by_age_gender <- pf %>% 
  filter(!is.na(gender)) %>%
  group_by(age, gender) %>% 
  summarise(friend_count_mean = mean(friend_count), friend_count_median = median(friend_count), n=n()) %>%    
  ungroup() %>%
  arrange(age)
head(pf.fc_by_age_gender, 20)
```

***

### Plotting Conditional Summaries
Notes:

```{r Plotting Conditional Summaries}

# Create a line graph showing the
# median friend count over the ages
# for each gender. Be sure to use
# the data frame you just created,
# pf.fc_by_age_gender.

# See the Instructor Notes for a hint.

# This assignment is not graded and
# will be marked as correct when you submit.

# ENTER YOUR CODE BELOW THIS LINE
# =================================================

library(ggplot2)
ggplot(aes(x = age, y = friend_count), 
              data = subset(pf, !is.na(gender))) +   geom_line(aes(color = gender), stat = 'summary', fun.y = median)
```

***

### Thinking in Ratios
Notes:

***

### Wide and Long Format
Notes:

***

### Reshaping Data
Notes:

```{r}
#install.packages('reshape2')
library(reshape2)

pf.fc_by_age_gender.wide <- dcast(pf.fc_by_age_gender, age~gender, value.var = 'friend_count_median')
```


***

### Ratio Plot
Notes:

```{r Ratio Plot}
# Plot the ratio of the female to male median
# friend counts using the data frame
# pf.fc_by_age_gender.wide.

# Think about what geom you should use.
# Add a horizontal line to the plot with
# a y intercept of 1, which will be the
# base line. Look up the documentation
# for geom_hline to do that. Use the parameter
# linetype in geom_hline to make the
# line dashed.

# The linetype parameter can take the values 0-6:
# 0 = blank, 1 = solid, 2 = dashed
# 3 = dotted, 4 = dotdash, 5 = longdash
# 6 = twodash

# This assignment is not graded and
# will be marked as correct when you submit.

# ENTER YOUR CODE BELOW THIS LINE
# =================================================

ggplot(aes(x = age, y = female/male), 
              data = pf.fc_by_age_gender.wide) +   
  geom_line() + geom_hline(aes(yintercept=1, color= 'red'),linetype = 2, alpha= .75)




```

***

### Third Quantitative Variable
Notes:

```{r Third Quantitative Variable}
# Create a variable called year_joined
# in the pf data frame using the variable
# tenure and 2014 as the reference year.

# The variable year joined should contain the year
# that a user joined facebook.

# See the Instructor Notes for three hints if you get
# stuck. Scroll down slowly to see one hint at a time
# if you would like some guidance.

# This programming exercise WILL BE automatically graded.

# DO NOT ALTER THE CODE BELOW THIS LINE
# ========================================================

pf$year_joined <- NA
pf$year_joined <- floor(2014 - pf$tenure/365)

```

***

### Cut a Variable
Notes:

```{r Cut a Variable}
# Create a new variable in the data frame
# called year_joined.bucket by using
# the cut function on the variable year_joined.

# You need to create the following buckets for the
# new variable, year_joined.bucket

#        (2004, 2009]
#        (2009, 2011]
#        (2011, 2012]
#        (2012, 2014]

# Note that a parenthesis means exclude the year and a
# bracket means include the year.

# Look up the documentation for cut or try the link
# in the Instructor Notes to accomplish this task.

# DO NOT DELETE THE TWO LINES OF CODE BELOW THIS LINE
# ========================================================================

pf$year_joined.bucket <- cut(pf$year_joined,breaks = c(2004,2009,2011,2012, 2014))
```

***

### Plotting it All Together
Notes:

```{r Plotting it All Together}
# Create a line graph of friend_count vs. age
# so that each year_joined.bucket is a line
# tracking the median user friend_count across
# age. This means you should have four different
# lines on your plot.

# You should subset the data to exclude the users
# whose year_joined.bucket is NA.

# If you need a hint, see the Instructor Notes.

# This assignment is not graded and
# will be marked as correct when you submit.

# ENTER YOUR CODE BELOW THIS LINE
# ===================================================
# Explore the variable
table(pf$year_joined.bucket, useNA = 'ifany')

#Plot
ggplot(aes(x = age, y = friend_count), 
              data = subset(pf, !is.na(year_joined.bucket))) + 
  geom_line(aes(color = year_joined.bucket), stat = 'summary', fun.y = median)

```
We confirm that people that has longer tenure ('(2004,2009]') has more friends.
***

### Plot the Grand Mean
Notes:

```{r Plot the Grand Mean}
# Write code to do the following:

# (1) Add another geom_line to code below
# to plot the grand mean of the friend count vs age.

# (2) Exclude any users whose year_joined.bucket is NA.

# (3) Use a different line type for the grand mean.

# As a reminder, the parameter linetype can take the values 0-6:

# 0 = blank, 1 = solid, 2 = dashed
# 3 = dotted, 4 = dotdash, 5 = longdash
# 6 = twodash

# This assignment is not graded and
# will be marked as correct when you submit.

# The code from the last programming exercise should
# be your starter code!

# ENTER YOUR CODE BELOW THIS LINE
# ==================================================================

ggplot(aes(x = age, y = friend_count), 
              data = subset(pf, !is.na(year_joined.bucket))) + 
  geom_line(aes(color = year_joined.bucket), stat = 'summary', fun.y = mean, linetype=1)+ geom_line(stat = 'summary', fun.y= mean, linetype=2, color='black', alpha=0.5) 
#last one is the grand_mean: the overall mean.
```

***

### Friending Rate
Notes: Calculate the firends per day of tenure. 

```{r Friending Rate}
#tenure of more than 1 day
with(subset(pf, tenure>=1), summary(friend_count / tenure)) # firends per day

```

***

### Friendships Initiated
Notes:

What is the median friend rate?

What is the maximum friend rate?

```{r Friendships Initiated}
# Create a line graph of mean of friendships_initiated per day (of tenure)
# vs. tenure colored by year_joined.bucket.

# You need to make use of the variables tenure,
# friendships_initiated, and year_joined.bucket.

# You also need to subset the data to only consider user with at least
# one day of tenure.

# This assignment is not graded and
# will be marked as correct when you submit.

# ENTER YOUR CODE BELOW THIS LINE
# ========================================================================


ggplot(aes(x = tenure, y = friendships_initiated / tenure), 
              data = subset(pf, tenure>=1)) + 
  geom_line(aes(color = year_joined.bucket), stat = 'summary', fun.y = mean, linetype=1)


```
People with more tenure typically initiate less friendships.

***

### Bias-Variance Tradeoff Revisited
Notes: Lot of noise in the previous graph. Set x axis = x*round(tenure/x), where the bigger the x, the smoother the graph.

```{r Bias-Variance Tradeoff Revisited}

ggplot(aes(x = tenure, y = friendships_initiated / tenure),
       data = subset(pf, tenure >= 1)) +
  geom_line(aes(color = year_joined.bucket),
            stat = 'summary',
            fun.y = mean)

ggplot(aes(x = 7 * round(tenure / 7), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = 30 * round(tenure / 30), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = 90 * round(tenure / 90), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

# Instead of geom_line(), use geom_smooth() to add a smoother to the plot.
# You can use the defaults for geom_smooth() but do color the line
# by year_joined.bucket

# ALTER THE CODE BELOW THIS LINE
# ==============================================================================

ggplot(aes(x = 7 * round(tenure / 7), y = friendships_initiated / tenure),data = subset(pf, tenure >= 0)) +
  geom_smooth(aes(color = year_joined.bucket))
```

***

### Sean's NFL Fan Sentiment Study
Notes:

***

## Yogurt

### Introducing the Yogurt Data Set
Notes:

***

### Histograms Revisited
Notes:

```{r Histograms Revisited}
yo <- read.csv('yogurt.csv')
str(yo)

# CHange id into a factor

yo$id <- factor(yo$id)
str(yo)

ggplot(aes(x=price ), data= yo)+ geom_histogram()

```

***

### Number of Purchases
Notes:

```{r Number of Purchases}
# Create a new variable called all.purchases,
# which gives the total counts of yogurt for
# each observation or household.

# One way to do this is using the transform
# function. You can look up the function transform
# and run the examples of code at the bottom of the
# documentation to figure out what it does.

# The transform function produces a data frame
# so if you use it then save the result to 'yo'!

# OR you can figure out another way to create the
# variable.

# DO NOT ALTER THE CODE BELOW THIS LINE
# ========================================================
yo <- read.csv('yogurt.csv')

yo$id <- factor(yo$id)


# ENTER YOUR CODE BELOW THIS LINE
# ========================================================

yo$all.purchases <- NA

yo <-  yo %>%
  mutate(all.purchases = strawberry+ blueberry+ pina.colada + plain +mixed.berry )

head(yo, 20)

#alternative solution

yo <- transform(yo, all.purchases = strawberry+ blueberry+ pina.colada + plain +mixed.berry)

summary(yo$all.purchases)

ggplot(aes(x=all.purchases ), data= yo)+ geom_histogram(binwidth = 1)

```
We can see most of houses buy 0-1 yogurts.
***

### Prices over Time
Notes:

```{r Prices over Time}
# Create a scatterplot of price vs time.

# This will be an example of a time series plot.

# Resolve overplotting issues by using
# techniques you learned in Lesson 4.

# What are some things that you notice?

# ENTER YOUR CODE BELOW THIS LINE
# ================================================
str(yo)

ggplot(aes(x=time/365, y =price), data= yo)+ geom_jitter(alpha=1/10, shape=21, color="blue")
```


Most common yogurt has a price that increases over time. Other prices may be related to using coupons.
***

### Sampling Observations
Notes:

***

### Looking at Samples of Households

```{r Looking at Sample of Households}
set.seed(4230)
set.seed(9)

sample.ids <- sample(levels(yo$id),16)

ggplot(aes(x=time/365, y =price), data= subset(yo, id %in% sample.ids)) + facet_wrap(~id)+ geom_line()+ geom_point(aes(size=all.purchases),alpha=1/3,color="blue")
```
We can se some houses do not buy much yogurt. Others buy a lot, or use coupons. Also, it can be seen that in general price increase over time.
***

### The Limits of Cross Sectional Data
Notes:

***

### Many Variables
Notes:

***

### Scatterplot Matrix
Notes:

***

### Even More Variables
Notes:

***

### Heat Maps
Notes:

```{r}
nci <- read.table("nci.tsv")
colnames(nci) <- c(1:64)
```

```{r}
nci.long.samp <- melt(as.matrix(nci[1:200,]))
names(nci.long.samp) <- c("gene", "case", "value")
head(nci.long.samp)

ggplot(aes(y = gene, x = case, fill = value),
  data = nci.long.samp) +
  geom_tile() +
  scale_fill_gradientn(colours = colorRampPalette(c("blue", "red"))(100))
```


***

### Analyzing Three of More Variables
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

