Lesson 3
========================================================

***

### What to Do First?
Notes:


```{r}
setwd('~/Git/data-analyst-nanodegree/P4/eda-course-materials/lesson3/')
getwd()

list.files()
pf <- read.csv('pseudo_facebook.tsv',sep = '\t')
```

### Pseudo-Facebook User Data
Notes:

```{r Pseudo-Facebook User Data}

```

***

### Histogram of Users' Birthdays
Notes:

```{r Histogram of Users\' Birthdays}
#install.packages('ggplot2')
library(ggplot2)
names(pf)

# Does not work in mac OSx
#qplot(x = dob_day, data = pf) +
#  scale_x_discrete(breaks = 1:31)

#install.packages('ggthemes', dependencies = TRUE) 
library(ggthemes) 

ggplot(aes(x = dob_day), data = pf ) +
  geom_histogram(binwidth = 1) +
  scale_x_continuous(breaks = 1:31)
```

***

#### What are some things that you notice about this histogram?
Response:

***

### Moira's Investigation
Notes:

***

### Estimating Your Audience Size
Notes:
```{r}
ggplot(data = pf, aes(x = dob_day)) + 
  geom_histogram(binwidth = 1) + 
   scale_x_continuous(breaks = 1:31) + 
   facet_wrap(~dob_month)
```

***

#### Think about a time when you posted a specific message or shared a photo on Facebook. What was it?
Response:

#### How many of your friends do you think saw that post?
Response:

#### Think about what percent of your friends on Facebook see any posts or comments that you make in a month. What percent do you think that is?
Response:

***

### Perceived Audience Size
Notes:

***
### Faceting
Notes:

```{r Faceting}

```

#### Let’s take another look at our plot. What stands out to you here?
Response:

***

### Be Skeptical - Outliers and Anomalies
Notes:

***

### Moira's Outlier
Notes:
#### Which case do you think applies to Moira’s outlier?
Response:

***

### Friend Count
Notes:

#### What code would you enter to create a histogram of friend counts?

```{r Friend Count}
qplot(x = friend_count, data = pf, xlim = c(0,1000))
```

#### How is this plot similar to Moira's first plot?
Response:

***

### Limiting the Axes
Notes:

```{r Limiting the Axes}

```

### Exploring with Bin Width
Notes:

***

### Adjusting the Bin Width
```{r}
qplot(x = friend_count, data = pf, xlim = c(0,1000), binwidth = 25)
```

Notes:

### Faceting Friend Count
```{r Faceting Friend Count}
# What code would you add to create a facet the histogram by gender?
# Add it to the code below.
qplot(x = friend_count, data = pf, binwidth = 10) +
  scale_x_continuous(limits = c(0, 750), breaks = seq(0, 750, 50))+ 
# Correct solution
  facet_wrap(~gender)

#My solution
qplot(x = friend_count, data = subset(pf, pf$gender =='female'), binwidth = 10) +
  scale_x_continuous(limits = c(0, 1000),
                     breaks = seq(0, 1000, 50))
#title("females")
                      
qplot(x = friend_count, data = subset(pf, pf$gender =='male'), binwidth = 10) +
  scale_x_continuous(limits = c(0, 1000),
                     breaks = seq(0, 1000, 50))+
#title("males")
                      

                      
```

***

### Omitting NA Values
Notes:

```{r Omitting NA Values}

```

***

### Statistics 'by' Gender
Notes:

```{r Statistics \'by\' Gender}
qplot(x = friend_count, data = subset(pf, !is.na(pf$gender)), binwidth = 10) +
  scale_x_continuous(limits = c(0, 750), breaks = seq(0, 750, 50))+ 
# Correct solution
  facet_wrap(~gender)
table(pf$gender)
by(pf$friend_count, pf$gender, summary )
```

#### Who on average has more friends: men or women?
Response:

#### What's the difference between the median friend count for women and men?
Response:

#### Why would the median be a better measure than the mean?
Response:

***

### Tenure
Notes:

```{r Tenure}
qplot(x=tenure, data = pf, color = I('black'), fill = I('#099DD9'), binwidth = 30)
```

***

#### How would you create a histogram of tenure by year?

```{r Tenure Histogram by Year}
qplot(x=tenure/365, data = pf, color = I('black'), fill = I('#F79420'), binwidth = .5) +
  scale_x_continuous(breaks = seq(1,7,1), limits = c(0,7))
```

***

### Labeling Plots
Notes:

```{r Labeling Plots}
qplot(x=tenure/365, data = pf, color = I('black'), fill = I('#F79420'), binwidth = .5,
      xlab = 'Number of years using Facebook', ylab= 'Number of users in sample')+
  scale_x_continuous(breaks = seq(0,7,1), limits = c(0,7))
```

***

### User Ages
Notes:

```{r User Ages}
qplot(x=age, data = subset(pf, !is.na(pf$gender)), color = I('black'), fill = I('#5760AB'), binwidth = 2,
      xlab = 'Age', ylab= 'Number of users in sample')+
  scale_x_continuous(breaks = seq(13,113,10), limits = c(13,113))+
  facet_wrap(~gender)
summary(pf$age)
```

#### What do you notice?
Response:

***

### The Spread of Memes
Notes:

***

### Lada's Money Bag Meme
Notes:

***

### Transforming Data
Notes:

***

### Add a Scaling Layer
Notes:

```{r Add a Scaling Layer}
#install.packages('gridExtra')
library('gridExtra')

# define individual plots
p1 <- qplot(x=friend_count, data = pf, color = I('black'), fill = I('#5760AB'), binwidth = 50,
      xlab = ' Number of friends', ylab= 'Number of users in sample')

p2 <- qplot(x=log10(friend_count+1), data = pf, color = I('black'), fill = I('#5760AB'), binwidth = 0.1,
      xlab = ' Number of friends', ylab= 'Number of users in sample') + scale_x_log10()
p3 <- qplot(x=sqrt(friend_count), data = pf, color = I('black'), fill = I('#5760AB'), binwidth = 2,
      xlab = ' Number of friends',  ylab= 'Number of users in sample')

# arrange plots in grid
grid.arrange(p1, p2, p3, ncol=1)

```

```{r Add a Scaling Layer: ggplot}
#install.packages('gridExtra')
library('gridExtra')

# define individual plots
p1 <- ggplot(aes(x = friend_count), data = pf, color = I('black'), fill = I('#5760AB'), binwidth = 50,
      xlab = ' Number of friends', ylab= 'Number of users in sample') +  geom_histogram()

p2 <- p1 + scale_x_log10()
p3 <- p1 + scale_x_sqrt()

# arrange plots in grid
grid.arrange(p1, p2, p3, ncol=1)

```
***


### Frequency Polygons

```{r Frequency Polygons}
qplot(x = friend_count, data = subset(pf, !is.na(pf$gender)), binwidth = 10, geom ='freqpoly', color = gender) +
  scale_x_continuous(limits = c(0, 750), breaks = seq(0, 750, 50))

ggplot(aes(x = friend_count, y = ..count../sum(..count..)), data = subset(pf, !is.na(gender))) + 
  geom_freqpoly(aes(color = gender), binwidth=10) + 
  scale_x_continuous(limits = c(0, 1000), breaks = seq(0, 1000, 50)) + 
  xlab('Friend Count') + 
  ylab('Percentage of users with that friend count')
       
       
ggplot(aes(x = www_likes, y = ..count../sum(..count..)), data = subset(pf, !is.na(gender))) + 
  geom_freqpoly(aes(color = gender)) + 
  scale_x_continuous() + scale_x_log10() + xlab('Friend Count') +  ylab('Percentage of users with that friend count')       
```

***

### Likes on the Web
Notes:

```{r Likes on the Web}
# likes received by males
sum(subset(pf$www_likes,pf$gender=="male"))

# likes received by females
sum(subset(pf$www_likes,pf$gender=="female"))


# SOLUTION

by(pf$www_likes,pf$gender,sum)
```


***

### Box Plots
Notes:

```{r Box Plots}
qplot( x= gender, y= friend_count,
       data= subset(pf, !is.na(gender) & pf$friend_count<=1000),
       geom = 'boxplot'
)
# another option coord_cartesian(ylim= c(0,1000))
```

#### Adjust the code to focus on users who have friend counts between 0 and 1000.

```{r}
qplot( x= gender, y= friendships_initiated,
       data= subset(pf, !is.na(gender) & pf$friend_count<=1000),
       geom = 'boxplot'
)+
  coord_cartesian(ylim = c(0,1000))

by(pf$friendships_initiated,pf$gender, summary)
```

***

### Box Plots, Quartiles, and Friendships
Notes:

```{r Box Plots, Quartiles, and Friendships}
summary(pf$mobile_likes)

summary(pf$mobile_likes > 0)

pf$mobile_check_in <- NA
pf$mobile_check_in <- ifelse(pf$mobile_likes>0,1,0)
pf$mobile_check_in <- factor(pf$mobile_check_in)

summary(pf$mobile_check_in)

#return percentage
sum( pf$mobile_check_in==1)/length(pf$mobile_check_in)
```

#### On average, who initiated more friendships in our sample: men or women?
Response:
#### Write about some ways that you can verify your answer.
Response:
```{r Friend Requests by Gender}

```

Response:

***

### Getting Logical
Notes:

```{r Getting Logical}

```

Response:

***

### Analyzing One Variable
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!