Lesson 4
========================================================

### Load data

```{r}
setwd('~/Git/data-analyst-nanodegree/P4/eda-course-materials/lesson4/') # ABSOLUTE PATH! CHANGE if PARENT directory changes!
getwd()

list.files()
pf <- read.csv('pseudo_facebook.tsv',sep = '\t')
```
***

### Scatterplots and Perceived Audience Size
Notes:

***

### Scatterplots
Notes:

```{r Scatterplots}
library(ggplot2)
qplot(age, friend_count, data=pf)
```

***

#### What are some things that you notice right away?
Response:

***

### ggplot Syntax
Notes:

```{r ggplot Syntax}
ggplot(aes(x=age, y=friend_count), data=pf) + geom_point() + xlim(13,90)

summary(pf$age)
```

***

### Overplotting
Notes:

```{r Overplotting}
ggplot(aes(x=age, y=friend_count), data=pf) + geom_jitter(alpha=1/20) + xlim(13,90) # add transparency and also dispace the points to see more clearly
```

#### What do you notice in the plot?
Response:

***

### Coord_trans()
Notes:

```{r Coord_trans()}
# y axis with sqrt transformation
ggplot(aes(x=age, y=friend_count), data=pf) + geom_point(alpha=1/20) + xlim(13,90)+ 
  coord_trans(y = "sqrt")## add transparency and also dispace the points to see more clearly

#Using jitter: add position_jitter in order to avoid negative sqrts
ggplot(aes(x=age, y=friend_count), data=pf) + geom_jitter(alpha=1/20, position = position_jitter(h=0)) + xlim(13,90)+ 
  coord_trans(y = "sqrt")## add transparency and also dispace the points to see more clearly



```

#### Look up the documentation for coord_trans() and add a layer to the plot that transforms friend_count using the square root function. Create your plot!

```{r}
# view previous
```

#### What do you notice?

***

### Alpha and Jitter
Notes:

```{r Alpha and Jitter}
#friends initiated and age
ggplot(aes(x=age, y=friendships_initiated), data=pf) + geom_jitter(alpha=1/50, position = position_jitter(h=0)) + xlim(13,90)+ 
  coord_trans(y = "sqrt")## add transparency and also dispace the points to see more clearly


```

***

### Overplotting and Domain Knowledge
Notes:

***

### Conditional Means
Notes:

```{r Conditional Means}
library(dplyr)

pf.fc_by_age <- pf %>%
  group_by(age) %>%
  summarise(friend_count_mean = mean(friend_count), friend_count_median = median(friend_count), n=n()) %>%    
  arrange(age)

head(pf.fc_by_age, 20)


# Plot mean friend count vs. age using a line graph.
# Be sure you use the correct variable names
# and the correct data frame. You should be working
# with the new data frame created from the dplyr
# functions. The data frame is called 'pf.fc_by_age'.

# Use geom_line() rather than geom_point to create
# the plot. You can look up the documentation for
# geom_line() to see what it does.

# ENTER ALL OF YOUR CODE TO CREATE THE PLOT BELOW THIS LINE.
# ===========================================================

ggplot(aes(x = age, y= friend_count_mean), data=pf.fc_by_age) + geom_line()
```




Create your plot!

```{r Conditional Means Plot}

```

***

### Overlaying Summaries with Raw Data
Notes:

```{r Overlaying Summaries with Raw Data}

ggplot(aes(x=age, y=friendships_initiated), data=pf) + geom_jitter(alpha=1/50, position = position_jitter(h=0), color= "orange") + xlim(13,90)+ 
  coord_trans(y = "sqrt") +## add transparency and also dispace the points to see more clearly
geom_line(stat = 'summary', fun.y=mean) +
geom_line(stat = 'summary', fun.y=quantile, fun.args= list(probs=.1), linetype = 2, color = 'blue')+
geom_line(stat = 'summary', fun.y=quantile, fun.args= list(probs=.5), linetype = 2, color = 'blue')+
geom_line(stat = 'summary', fun.y=quantile, fun.args= list(probs=.9), linetype = 2, color = 'blue')  
```

#### What are some of your observations of the plot?
Response:

***

### Moira: Histogram Summary and Scatterplot
See the Instructor Notes of this video to download Moira's paper on perceived audience size and to see the final plot.

Notes:

***

### Correlation
Notes:

```{r Correlation}
cor(pf$age,pf$friend_count)
```

Look up the documentation for the cor.test function.

What's the correlation between age and friend count? Round to three decimal places.
Response:

***

### Correlation on Subsets
Notes:

```{r Correlation on Subsets}
#Caculate correlation for ages <= 70
with(subset(pf, age<= 70), cor.test(age, friend_count))
# Now relation is stronger
```

***

### Correlation Methods
Notes:

***

## Create Scatterplots
Notes:

```{r}

# Create a scatterplot of likes_received (y)
# vs. www_likes_received (x). Use any of the
# techniques that you've learned so far to
# modify the plot.

# ENTER ALL OF YOUR CODE TO CREATE THE PLOT BELOW THIS LINE.
# ===========================================================

ggplot(aes(x=likes_received, y= www_likes_received),data=pf)+ geom_point(alpha= 1/5, position = position_jitter(h=0))
```

***

### Strong Correlations
Notes:

```{r Strong Correlations}
ggplot(aes(x=likes_received, y= www_likes_received),data=pf)+ geom_point(alpha= 1/20, position = position_jitter(h=0))+
xlim(0,quantile(pf$likes_received,0.95)) +
ylim(0,quantile(pf$www_likes_received,0.95))+
  geom_smooth(method = 'lm', color='red')

with(pf, cor.test(likes_received, www_likes_received))
```

What's the correlation betwen the two variables? Include the top 5% of values for the variable in the calculation and round to 3 decimal places.

```{r Correlation Calcuation}

```

Response:

***

### Moira on Correlation
Notes:

***

### More Caution with Correlation
Notes:

```{r More Caution With Correlation}
#install.packages('alr3')
library(alr3)

data(Mitchell)
# Create a scatterplot of temperature (Temp)
# vs. months (Month).

# ENTER ALL OF YOUR CODE TO CREATE THE PLOT BELOW THIS LINE.
# ===========================================================
ggplot(aes(y=Temp, x=Month), data=Mitchell) + geom_point(alpha=1/3)
with(Mitchell,  cor.test(Mitchell$Month,Mitchell$Temp))
```

Create your plot!

```{r Temp vs Month}
Mitchell$new_Month <- NA
Mitchell$new_Month <- Mitchell$Month %% 12 # Convert to numeric month instead of periodic count

ggplot(aes(y=Temp, x=new_Month), data=Mitchell) + geom_point(alpha=1/5)
with(Mitchell,  cor.test(Mitchell$new_Month,Mitchell$Temp))

# Preserving pattern over time
ggplot(aes(y=Temp, x=Month), data=Mitchell) + geom_point(alpha=1/5) + scale_x_discrete(breaks= c(0, nrow(Mitchell),12)) # when squeezed to have low height we can observe a sinusoidal pattern
```

***

### Noisy Scatterplots
a. Take a guess for the correlation coefficient for the scatterplot.

b. What is the actual correlation of the two variables?
(Round to the thousandths place)

```{r Noisy Scatterplots}
pf$age_with_months <- NA
pf$age_with_months <- with(pf, age + (12 - dob_month / 12))


ggplot(aes(x = age_with_months, y= friend_count), data=pf) + geom_line()
```

***

### Making Sense of Data
Notes:

```{r Making Sense of Data}



```

***

### A New Perspective

What do you notice?
Response:

Watch the solution video and check out the Instructor Notes!
Notes:

***

### Understanding Noise: Age to Age Months
Notes:

```{r Understanding Noise: Age to Age Months}
pf.fc_by_age_months <- pf %>%
  group_by(age_with_months) %>%
  summarise(friend_count_mean = mean(friend_count),
            friend_count_median = median(friend_count), 
            n = n()) %>%
  arrange(age_with_months) 

head(pf.fc_by_age)

```

***

### Age with Months Means

```{r Age with Months Means}
# Create a new scatterplot showing friend_count_mean
# versus the new variable, age_with_months. Be sure to use
# the correct data frame (the one you create in the last
# exercise) AND subset the data to investigate
# users with ages less than 71.

# ENTER YOUR CODE BELOW THIS LINE.
# ===============================================================


ggplot(aes(x = age_with_months, y= friend_count_mean), data=subset(pf.fc_by_age_months, age_with_months<71)) + geom_line()
```

Programming Assignment
```{r Programming Assignment}

```

***

### Noise in Conditional Means

```{r Noise in Conditional Means}
p1 <- ggplot(aes(x = age, y= friend_count_mean), data=subset(pf.fc_by_age, age<71)) + geom_line()+ geom_smooth() 

p2 <- ggplot(aes(x = age_with_months, y= friend_count_mean), data=subset(pf.fc_by_age_months, age_with_months<71)) + geom_line() + geom_smooth() 


p3 <- ggplot(aes(x = round(age/5)*5, y= friend_count_mean), data=subset(pf.fc_by_age, age<71)) + geom_line(stat= 'summary', fun.y= mean)+ geom_smooth() 





library(gridExtra)

grid.arrange(p1,p2,p3, ncol=1)


```

***

### Smoothing Conditional Means
Notes:

```{r Smoothing Conditional Means}

```

***

### Which Plot to Choose?
Notes:

***

### Analyzing Two Variables
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

