---
title: 'Project 2 : Explartion of Red Wine'
author: "Isabel María Villalba Jiménez"
date: "6 de diciembre de 2016"
output:
  html_document:
    depth: 1
    fig_caption: yes
    fig_height: 6
    fig_width: 8
    toc: yes
  pdf_document:
    citation_package: natbib
    fig_caption: yes
    fig_height: 6
    fig_width: 7
    highlight: tango
    keep_tex: yes
    latex_engine: pdflatex
    toc: no
bibliography: P4.bib
biblio-style: apsr
---


```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using
# in your analysis in this code chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk.
# This prevents the code from displaying in the knitted HTML output.
# You should set echo=FALSE for all code chunks in your file.

library(ggplot2)
```

```{r echo=FALSE, Load_the_Data}
# Load the Data
redwines <- read.csv('wineQualityReds.csv')
names(redwines)
summary(redwines)


  
# New variables 
redwines$quality.factor <- factor(redwines$quality)


redwines$quality.cat <- NA
redwines$quality.cat <- ifelse(redwines$quality>=7, 'good',   'medium')
redwines$quality.cat <- ifelse(redwines$quality<=4, 'bad',  redwines$quality.cat) # if not, leave the previous values
redwines$quality.cat <- factor(redwines$quality.cat, , levels = list('bad', 'medium','good')) # set the order!

str(redwines) #summary of values for each variable

unique(redwines$quality.cat)
```

# Univariate Plots Section
```{r echo=FALSE, Univariate_Plots}
ggplot(aes(x = quality), data = redwines) + geom_histogram(binwidth = 1) + ggtitle('Quality')
 
ggplot(aes(x = sulphates), data = redwines) + geom_histogram(binwidth = .02) + scale_x_log10() + ggtitle('Sulphates')

library(GGally)
ggpairs(redwines, axisLabels = 'internal', lower = list(continous = 'points', mapping = aes(shape= quality.cat, color=quality.cat), size = 8), upper = list(continous = 'smooth', discrete = 'facetwrap', size = 8), columns=2:10) + theme_light(base_size = 8)
```

## Fixed and volatile acidity
```{r echo=FALSE, Univariate_Plots_acidity}
library(gridExtra)
#Fixed
p1_a <- ggplot(aes(x = fixed.acidity, fill= quality.cat), data = redwines) + geom_histogram(binwidth=.5)# +  scale_x_log10() #+ theme_light(base_size = 10)
p1_b <-ggplot(aes(y = fixed.acidity, x= quality.cat, color= quality.cat), data = redwines) + geom_boxplot()


# Volatile
p2_a <-ggplot(aes(x = volatile.acidity, fill= quality.cat ), data = redwines) + geom_histogram(binwidth = .05)+  scale_x_log10() #+ theme_light(base_size = 10)
p2_b <-ggplot(aes(y = volatile.acidity, x= quality.cat, color= quality.cat), data = redwines) + geom_boxplot()

# Citric acid
p3_a <-ggplot(aes(x = citric.acid, fill= quality.cat ), data = redwines) + geom_histogram(binwidth = .05)#+ theme_light(base_size = 10)
p3_b <-ggplot(aes(y = citric.acid, x= quality.cat, color= quality.cat), data = redwines) + geom_boxplot()


grid.arrange(p1_a, p1_b,p2_a,p2_b, p3_a,p3_b, ncol=2) + theme_light(base_size = 10)
```
```{r echo=FALSE, Univariate_Plots_residual_sugar}
#Residual sugar
p4_a <- ggplot(aes(x = residual.sugar, fill= quality.cat), data = redwines) + geom_histogram(binwidth=.1) +  scale_x_log10() #+ theme_light(base_size = 10)
p4_b <-ggplot(aes(y = residual.sugar, x= quality.cat, color= quality.cat), data = redwines) + geom_boxplot() + coord_trans(y='log10') #transformed y

grid.arrange(p4_a,p4_b, ncol=2) + theme_light(base_size = 10)
```
```{r echo=FALSE, Univariate_Plots_chlorides}
#Chlorides
p5_a <- ggplot(aes(x = chlorides, fill= quality.cat), data = redwines) + geom_histogram(binwidth=.1) +  scale_x_log10() #+ theme_light(base_size = 10)
p5_b <-ggplot(aes(y = chlorides, x= quality.cat, color= quality.cat), data = redwines) + geom_boxplot() + coord_trans(y='log10') #transformed y

#Free sulfur dioxide
p6_a <- ggplot(aes(x = free.sulfur.dioxide, fill= quality.cat), data = redwines) + geom_histogram(binwidth=.1) +  scale_x_log10() #+ theme_light(base_size = 10)
p6_b <-ggplot(aes(y = free.sulfur.dioxide, x= quality.cat, color= quality.cat), data = redwines) + geom_boxplot() + coord_trans(y='log10') #transformed y

#Total sulfur dioxide
p7_a <- ggplot(aes(x = total.sulfur.dioxide, fill= quality.cat), data = redwines) + geom_histogram(binwidth=.1) +  scale_x_log10() #+ theme_light(base_size = 10)
p7_b <-ggplot(aes(y = total.sulfur.dioxide, x= quality.cat, color= quality.cat), data = redwines) + geom_boxplot() + coord_trans(y='log10') #transformed y


grid.arrange(p5_a,p5_b,p6_a,p6_b,p7_a,p7_b, ncol=2) + theme_light(base_size = 10)
```
```{r echo=FALSE, Univariate_Plots_density}
#Density
p8_a <- ggplot(aes(x = density, fill= quality.cat), data = redwines) + geom_histogram(binwidth=.0005) +  scale_x_log10() #+ theme_light(base_size = 10)
p8_b <-ggplot(aes(y = density, x= quality.cat, color= quality.cat), data = redwines) + geom_boxplot() + coord_trans(y='log10') #transformed y

#pH
p9_a <- ggplot(aes(x = pH, fill= quality.cat), data = redwines) + geom_histogram(binwidth=.01) +  scale_x_log10() #+ theme_light(base_size = 10)
p9_b <-ggplot(aes(y = pH, x= quality.cat, color= quality.cat), data = redwines) + geom_boxplot() + coord_trans(y='log10') #transformed y

#Sulphates
p10_a <- ggplot(aes(x = sulphates, fill= quality.cat), data = redwines) + geom_histogram(binwidth=.05) +  scale_x_log10() #+ theme_light(base_size = 10)
p10_b <-ggplot(aes(y = sulphates, x= quality.cat, color= quality.cat), data = redwines) + geom_boxplot() + coord_trans(y='log10') #transformed y

#Sulphates
p11_a <- ggplot(aes(x = alcohol, fill= quality.cat), data = redwines) + geom_histogram(binwidth=.01) +  scale_x_log10() #+ theme_light(base_size = 10)
p11_b <-ggplot(aes(y = alcohol, x= quality.cat, color= quality.cat), data = redwines) + geom_boxplot() + coord_trans(y='log10') #transformed y

grid.arrange(p8_a,p8_b,p9_a,p9_b,p10_a,p10_b,p11_a,p11_b, ncol=2) + theme_light(base_size = 10)
```
```{r echo=FALSE, Univariate_Plots_1}
library(gridExtra)
p1 <-ggplot(aes(x = fixed.acidity ), data = redwines) + geom_histogram(binwidth=.5)# +  scale_x_log10() #+ theme_light(base_size = 10)

p2 <-ggplot(aes(x = volatile.acidity ), data = redwines) + geom_histogram(binwidth = .05)#+  scale_x_log10() #+ theme_light(base_size = 10)

p3 <-ggplot(aes(x = citric.acid ), data = redwines) + geom_histogram(binwidth = .05)#+ theme_light(base_size = 10)

p4 <-ggplot(aes(x = residual.sugar), data = redwines) + geom_histogram(binwidth = .5) #+  scale_x_log10() #+ theme_light(base_size = 10)

p5 <-ggplot(aes(x = chlorides ), data = redwines) + geom_histogram(binwidth = .01)  #+  scale_x_log10()#+ theme_light

p6 <-ggplot(aes(x = free.sulfur.dioxide), data = redwines) + geom_histogram(binwidth = 5) #+  scale_x_log10() #+ theme_light(base_size = 10)

p7 <-ggplot(aes(x = total.sulfur.dioxide), data = redwines) + geom_histogram(binwidth = 10) #+ theme_light(base_size = 10)

p8 <-ggplot(aes(x = density), data = redwines) + geom_histogram(binwidth = .0005) #+ theme_light(base_size = 10)

p9 <-ggplot(aes(x = pH ), data = redwines) + geom_histogram(binwidth = .05)  #+  scale_x_log10()#+ theme_light

p10 <-ggplot(aes(x = sulphates), data = redwines) + geom_histogram(binwidth = 0.05) #+  scale_x_log10() #+ theme_light(base_size = 10)

p11 <-ggplot(aes(x = alcohol), data = redwines) + geom_histogram(binwidth = .5) #+ theme_light(base_size = 10)

p12 <-ggplot(aes(x = quality), data = redwines) + geom_bar() #+ theme_light(base_size = 10)

grid.arrange(p1,p2,p3,p4,p5,p6,p7,p8,p9,p10,p11,p12, ncol=4) + theme_light(base_size = 10)
```


# Univariate Analysis

### What is the structure of your dataset?

### What is/are the main feature(s) of interest in your dataset?

### What other features in the dataset do you think will help support your investigation into your feature(s) of interest?

### Did you create any new variables from existing variables in the dataset?

### Of the features you investigated, were there any unusual distributions? Did you perform any operations on the data to tidy, adjust, or change the form of the data? If so, why did you do this?



# Bivariate Plots Section
```{r echo=FALSE, Bivariate_Plots}

```

# Bivariate Analysis

### Talk about some of the relationships you observed in this part of the investigation. How did the feature(s) of interest vary with other features in the dataset?

### Did you observe any interesting relationships between the other features (not the main feature(s) of interest)?

### What was the strongest relationship you found?




# Multivariate Plots Section

```{r echo=FALSE, Multivariate_Plots}

```

# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the investigation. Were there features that strengthened each other in terms of looking at your feature(s) of interest?

### Were there any interesting or surprising interactions between features?

### OPTIONAL: Did you create any models with your dataset? Discuss the strengths and limitations of your model.

------

# Final Plots and Summary

### Plot One
```{r echo=FALSE, Plot_One}

```

### Description One


### Plot Two
```{r echo=FALSE, Plot_Two}

```

### Description Two


### Plot Three
```{r echo=FALSE, Plot_Three}

```

### Description Three

------

# Reflection



# Others 
```{r echo=FALSE, Univariate_Plots_log}
library(gridExtra)
p1 <-ggplot(aes(x = fixed.acidity ), data = redwines) + geom_histogram(binwidth = .01) +  scale_x_log10() #+ theme_light(base_size = 10)

p2 <-ggplot(aes(x = volatile.acidity ), data = redwines) + geom_histogram(binwidth = .01)+  scale_x_log10() #+ theme_light(base_size = 10)

p3 <-ggplot(aes(x = citric.acid ), data = redwines) + geom_histogram(binwidth = .01)#+ theme_light(base_size = 10)

p4 <-ggplot(aes(x = residual.sugar), data = redwines) + geom_histogram(binwidth = .05) +  scale_x_log10() #+ theme_light(base_size = 10)

p5 <-ggplot(aes(x = chlorides ), data = redwines) + geom_histogram(binwidth = .01)  +  scale_x_log10()#+ theme_light

p6 <-ggplot(aes(x = free.sulfur.dioxide), data = redwines) + geom_histogram(binwidth = 0.05) +  scale_x_log10() #+ theme_light(base_size = 10)

p7 <-ggplot(aes(x = total.sulfur.dioxide), data = redwines) + geom_histogram(binwidth = 0.05) #+ theme_light(base_size = 10)

p8 <-ggplot(aes(x = density), data = redwines) + geom_histogram(binwidth = .0005) #+ theme_light(base_size = 10)



grid.arrange(p1,p2,p3,p4,p5,p6,p7,p8, ncol=4) + theme_light(base_size = 10)
```
