---
title: 'Project 4  : Explartion of Red Wine Quality'
author: "Isabel María Villalba Jiménez"
date: "December 6th, 2016"
output:
  html_document:
    depth: 1
    fig_caption: yes
    fig_height: 6
    fig_width: 8
    #toc: yes
  pdf_document:
    citation_package: natbib
    fig_caption: yes
    fig_height: 6
    fig_width: 7
    highlight: tango
    keep_tex: yes
bibliography: P4.bib
biblio-style: apsr
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=4, fig.height=4, fig.path='Figs/',
                      echo=FALSE, warning=FALSE, message=FALSE, include = TRUE, results= 'hide')
```

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using
# in your analysis in this code chunk.


# Notice that the parameter "echo" was set to FALSE for this code chunk.
# This prevents the code from displaying in the knitted HTML output.
# You should set echo=FALSE for all code chunks in your file.
this_dir <- 'Git/data-analyst-nanodegree/P4-explore-and-summarize-data/' 
setwd("~")
setwd(this_dir)

library(ggplot2)
library(gridExtra)
```


#Analysis

In this work it will be analyzed the impact in quality of several parameters describing red wine. The dataset is curated by Udacity and comes from UCI repository [https://archive.ics.uci.edu/ml/datasets/Wine+Quality](https://archive.ics.uci.edu/ml/datasets/Wine+Quality) and consists of 1599 sample data for Red wine [https://docs.google.com/document/d/1qEcwltBMlRYZT-l699-71TzInWfk4W9q5rTCSvDVMpc/pub?embedded=true](https://docs.google.com/document/d/1qEcwltBMlRYZT-l699-71TzInWfk4W9q5rTCSvDVMpc/pub?embedded=true).

 In @cortez2009modeling it is shown that the most imporant features for assessing Red Wine quality are:
 
- sulphates

- pH

- total sulfur dioxide

# Variable summary


```{r echo=TRUE, results='markup', Load_the_Data}
# Load the Data
redwines <- read.csv('wineQualityReds.csv')

dim(redwines)
#names(redwines)
summary(redwines)

# New variables 
#redwines$quality.factor <- factor(redwines$quality)

redwines$quality.cat <- NA
redwines$quality.cat <- ifelse(redwines$quality>=7, 'good','medium')
redwines$quality.cat <- ifelse(redwines$quality<=4, 'bad',redwines$quality.cat) # if not, leave the previous values
redwines$quality.cat <- factor(redwines$quality.cat, levels = list('bad', 'medium','good')) # set the order!

print("Variables after dividing into quality groups")
str(redwines) #summary of values for each variable

#unique(redwines$quality.cat)
```

## Univariate Plots Section

In this section it will be analyzed each of the variables describing the wines.

###Quality

> The distribution of wine shows that most of wines have a quality between 5-6 points.

```{r echo=FALSE, Univariate_Plots,fig.width=3, fig.height=3}
ggplot(aes(x = quality), data = redwines) + geom_histogram(binwidth = 1, fill = "brown", color= "black") + ggtitle('Quality') +theme_light(base_size = 10)
 
#ggplot(aes(x = sulphates), data = redwines) + geom_histogram(binwidth = .02) + scale_x_log10() + ggtitle('Sulphates')

library(GGally)
#ggpairs(redwines, axisLabels = 'internal', lower = list(continous = 'points', mapping = aes(shape= quality.cat, color=quality.cat), size = 8), upper = list(continous = 'smooth', discrete = 'facetwrap', size = 8), columns=2:10) + theme_light(base_size = 8)
```

### Fixed and volatile acidity

> In the next plot it is deduced that the quality of the wine is directly proportional to the fixed acidity and acid levels and inversely proportional to volatile acidity. 

```{r echo=FALSE, Univariate_Plots_acidity,fig.width=9, fig.height=3}
library(gridExtra)
#Fixed
p1 <- ggplot(aes(x = fixed.acidity), data = redwines) + geom_histogram(binwidth=.5, color = "black", fill = "turquoise3") + ggtitle('Fixed acidity') +  theme_light(base_size = 10) # +  scale_x_log10() #+

# Volatile
p2 <-ggplot(aes(x = volatile.acidity), data = redwines) + geom_histogram(binwidth = .05,  color = "black", fill = "turquoise3")+  scale_x_log10() + ggtitle('Volatile acidity')  + theme_light(base_size = 10)

# Citric acid
p3 <-ggplot(aes(x = citric.acid), data = redwines) + geom_histogram(binwidth = .05,  color = "black", fill = "turquoise3") + ggtitle('Citric acidity') + theme_light(base_size = 10)


grid.arrange(p1,p2, p3, ncol=3) + theme_light(base_size = 10)
```

### Residual sugar

> The plot of residual suggar shows that the better the wine the higher the residual sugar levels. 

```{r echo=FALSE, Univariate_Plots_residual_sugar,fig.width=3, fig.height=3}
#Residual sugar
p4 <- ggplot(aes(x = residual.sugar), data = redwines) + geom_histogram(binwidth=.1,  color = "black", fill = "turquoise3") +  scale_x_log10() + ggtitle('Residual sugar')  + theme_light(base_size = 10)

p4
```

### Chlorides and sulfur dioxide

> In the plot of the chlorides, it can be onserved that the better the wine, the lower the chloride levels. For the sulfur dioxide, either the free or the total sulfur dioxide, high levels are indicator of medium quality, whereas bad and good wine have the same low amount of sulfure dioxide.

```{r echo=FALSE, Univariate_Plots_chlorides,fig.width=9, fig.height=3}
#Chlorides
p5 <- ggplot(aes(x = chlorides), data = redwines) + geom_histogram(binwidth=.1,  color = "black", fill = "turquoise3") +  scale_x_log10() + ggtitle('Chlorides')+ theme_light(base_size = 10)

#Free sulfur dioxide
p6 <- ggplot(aes(x = free.sulfur.dioxide), data = redwines) + geom_histogram(binwidth=.1,  color = "black", fill = "turquoise3") +  scale_x_log10() + ggtitle('Free sulfur dioxide') + theme_light(base_size = 10)

#Total sulfur dioxide
p7 <- ggplot(aes(x = total.sulfur.dioxide), data = redwines) + geom_histogram(binwidth=.1,  color = "black", fill = "turquoise3") +  scale_x_log10() + ggtitle('Total sulfur dioxide') + theme_light(base_size = 10)

grid.arrange(p5, p6, p7, ncol=3) + theme_light(base_size = 10)
```


### Density, pH, sulphates and alcohol

> The next plot shows that, generally, the lower the density, the better the quality of the wine. ALso, low pH levels are sign of beter quality. The higher the sulphates level, the better quality of the wine and also, good wines have higher amount of alcohol.

```{r echo=FALSE, Univariate_Plots_density, fig.width=6, fig.height=6}
#Density
p8 <- ggplot(aes(x = density), data = redwines) + geom_histogram(binwidth=.0005,  color = "black", fill = "turquoise3") +  scale_x_log10() + ggtitle('Density')+ theme_light(base_size = 10)

#pH
p9 <- ggplot(aes(x = pH), data = redwines) + geom_histogram(binwidth=.01, color = "black", fill = "turquoise3") +  scale_x_log10() + ggtitle('pH') + theme_light(base_size = 10)

#Sulphates
p10 <- ggplot(aes(x = sulphates), data = redwines) + geom_histogram(binwidth=.05,  color = "black", fill = "turquoise3") +  scale_x_log10() + ggtitle('Sulphates')+ theme_light(base_size = 10) 

#Alcohol
p11 <- ggplot(aes(x = alcohol), data = redwines) + geom_histogram(binwidth=.01, color = "black", fill = "turquoise3") +  scale_x_log10() + ggtitle('Alcohol')+ theme_light(base_size = 10)

grid.arrange(p8, p9, p10, p11, ncol=2) + theme_light(base_size = 10);
```



# Univariate Analysis

### What is the structure of your dataset?
```{r}
summary(redwines)
colnames(redwines)
```

There are 1599 red wines with 12 features (fixed.acidity, volatile.acidity, citric.acid, residual.sugar, chlorides, free.sulfur.dioxide, total.sulfur.dioxide, density, pH, sulphates, alcohol and quality). The variables quality is converted in a factor variable (adding a new variable named ```quality.cat```) with the following levels:

(worst) —————-> (best)

```quality.cat```: BAD (quality [0,4]), MEDIUM (quality (4,7)), GOOD (quality [7,10]),

Other observations:

The mean quality of the red wines is 5.636 and the median is 6. Q1 corresponds to 5 and Q3 to 6, hence, 50% of the data lays within 5-6 range of quality, this is the level MEDIUM.


### What is/are the main feature(s) of interest in your dataset?

The main feature of interest in this dataset is the **quality** of the wine.

### What other features in the dataset do you think will help support your investigation into your feature(s) of interest?

In @cortez2009modeling it is shown that the most imporant features for assessing Red Wine quality are: **sulphates**, **pH** and **total sulfur dioxide**.

### Did you create any new variables from existing variables in the dataset?

The amount of information available is enough to assess the quality of the wine and I did not create any new variables to support the analysis.

### Of the features you investigated, were there any unusual distributions? Did you perform any operations on the data to tidy, adjust, or change the form of the data? If so, why did you do this?

Most of the features had a normal distribution. Some of the features had quite skewed distributions and  many oultiers and I performed a log10 tranformation in order to have a better view.

```{r univariate_tranformation, fig.width=6, fig.height=6}
#Residual sugar
p4_base <- ggplot(aes(x = residual.sugar), data = redwines) 
p4_log <- p4_base + geom_histogram(binwidth=.1,  color = "black", fill = "turquoise3") +  scale_x_log10(breaks = c(1, 2, 5, 10)) + ggtitle('Residual sugar: log10 scaled')+ theme_light(base_size = 10)

p4_normal <- p4_base + geom_histogram(binwidth=.5,  color = "black", fill = "turquoise3") + ggtitle('Residual sugar')+ theme_light(base_size = 10) 

#Chlorides
p5_base <- ggplot(aes(x = chlorides), data = redwines) 
p5_log <- p5_base + geom_histogram(binwidth=.1,  color = "black", fill = "turquoise3") +  scale_x_log10(breaks = c(.05, .1 ,.2, .5)) + ggtitle('Chlorides: log10 scaled')+ theme_light(base_size = 8)

p5_normal <- p5_base + geom_histogram(binwidth=.02,  color = "black", fill = "turquoise3") + ggtitle('Chlorides')+ theme_light(base_size = 10) 

grid.arrange(p4_normal, p4_log, p5_normal, p5_log, ncol=2) + theme_light(base_size = 10)
```



# Bivariate Plots Section
```{r echo=FALSE, Bivariate_Plots}

```
### Fixed and volatile acidity

> In the next plot it is deduced that the quality of the wine is directly proportional to the fixed acidity and acid levels and inversely proportional to volatile acidity. 


```{r echo=FALSE, message= FALSE, Bivariate_Plots_acidity,fig.width=8, fig.height=6}
#Fixed acidity
p1_a <- ggplot(aes(x = fixed.acidity, fill= quality.cat), data = redwines) + geom_histogram(binwidth=.5, color = "black", size= 0.25) + ggtitle('Fixed acidity') + theme_light(base_size = 10) # +  scale_x_log10() #+ theme_light(base_size = 10)
p1_b <-ggplot(aes(y = fixed.acidity, x= quality.cat, fill= quality.cat), data = redwines) + geom_boxplot(outlier.alpha = 1/10) + geom_boxplot() + ggtitle('Fixed acidity') +  theme_light(base_size = 10)


# Volatile
p2_a <-ggplot(aes(x = volatile.acidity, fill= quality.cat ), data = redwines) + geom_histogram(binwidth = .05, color = "black", size= 0.25)+  scale_x_log10() + ggtitle('Volatile acidity')  + theme_light(base_size = 10)
p2_b <-ggplot(aes(y = volatile.acidity, x= quality.cat, fill= quality.cat), data = redwines) + geom_boxplot(outlier.alpha = 1/10) + geom_boxplot() + ggtitle('Volatile acidity') +  scale_y_log10() +  theme_light(base_size = 10)

# Citric acid
p3_a <-ggplot(aes(x = citric.acid, fill= quality.cat ), data = redwines) + geom_histogram(binwidth = .05, color = "black", size= 0.25) + ggtitle('Citric acidity') + theme_light(base_size = 10)
p3_b <-ggplot(aes(y = citric.acid, x= quality.cat, fill= quality.cat), data = redwines) + geom_boxplot(outlier.alpha = 1/10) + geom_boxplot() + ggtitle('Citric acidity') +  theme_light(base_size = 10)


grid.arrange(p1_a, p1_b, p2_a,p2_b, p3_a,p3_b, ncol=2) + theme_light(base_size = 10)
```

### Residual sugar

> The plot of residual suggar shows that the better the wine the higher the residual sugar levels. 

```{r echo=FALSE, Bivariate_Plots_residual_sugar,fig.width=8, fig.height=2}
#Residual sugar
p4_a <- ggplot(aes(x = residual.sugar, fill= quality.cat), data = redwines) + geom_histogram(binwidth=.1, color = "black", size= 0.25) +  scale_x_log10() + ggtitle('Residual sugar')  + theme_light(base_size = 10)
p4_b <-ggplot(aes(y = residual.sugar, x= quality.cat, fill= quality.cat), data = redwines) + geom_boxplot(outlier.alpha = 1/10) + scale_y_log10() + ggtitle('Residual sugar') +  theme_light(base_size = 10) #transformed y

grid.arrange(p4_a,p4_b, ncol=2) + theme_light(base_size = 10)
```

### Chlorides and sulfur dioxide

> In the plot of the chlorides, it can be onserved that the better the wine, the lower the chloride levels. For the sulfur dioxide, either the free or the total sulfur dioxide, high levels are indicator of medium quality, whereas bad and good wine have the same low amount of sulfure dioxide.

```{r echo=FALSE, Bivariate_Plots_chlorides,fig.width=8, fig.height=6}
#Chlorides
p5_a <- ggplot(aes(x = chlorides, fill= quality.cat), data = redwines) + geom_histogram(binwidth=.1, color = "black", size= 0.25) +  scale_x_log10() + ggtitle('Chlorides')+ theme_light(base_size = 10)
p5_b <-ggplot(aes(y = chlorides, x= quality.cat, fill= quality.cat), data = redwines) + geom_boxplot(outlier.alpha = 1/10) + scale_y_log10() + ggtitle('Chlorides') + theme_light(base_size = 10)#transformed y

#Free sulfur dioxide
p6_a <- ggplot(aes(x = free.sulfur.dioxide, fill= quality.cat), data = redwines) + geom_histogram(binwidth=.1, color = "black", size= 0.25) +  scale_x_log10() + ggtitle('Free sulfur dioxide') + theme_light(base_size = 10)
p6_b <-ggplot(aes(y = free.sulfur.dioxide, x= quality.cat, fill= quality.cat), data = redwines) + geom_boxplot(outlier.alpha = 1/10) + scale_y_log10() + ggtitle('Free sulfur dioxide') +  theme_light(base_size = 10) #transformed y

#Total sulfur dioxide
p7_a <- ggplot(aes(x = total.sulfur.dioxide, fill= quality.cat), data = redwines) + geom_histogram(binwidth=.1, color = "black", size= 0.25) +  scale_x_log10() + ggtitle('Total sulfur dioxide') + theme_light(base_size = 10)
p7_b <-ggplot(aes(y = total.sulfur.dioxide, x= quality.cat, fill= quality.cat), data = redwines) + geom_boxplot(outlier.alpha = 1/10) + scale_y_log10() + ggtitle('Total sulfur dioxide') +  theme_light(base_size = 10)#transformed y

grid.arrange(p5_a,p5_b,p6_a,p6_b,p7_a,p7_b, ncol=2) + theme_light(base_size = 10)
```


### Density, pH, sulphates and alcohol

> The next plot shows that, generally, the lower the density, the better the quality of the wine. ALso, low pH levels are sign of beter quality. The higher the sulphates level, the better quality of the wine and also, good wines have higher amount of alcohol.

```{r echo=FALSE, Bivariate_Plots_density, fig.width=8, fig.height=8}

#Density
p8_a <- ggplot(aes(x = density, fill= quality.cat), data = redwines) + geom_histogram(binwidth=.0005, color = "black", size= 0.25) +  scale_x_log10() + ggtitle('Density')+ theme_light(base_size = 10)
p8_b <-ggplot(aes(y = density, x= quality.cat, fill= quality.cat), data = redwines) + geom_boxplot(alpha= 1, outlier.alpha = 1/10) + scale_y_log10() + ggtitle('Density')+  theme_light(base_size = 10)#transformed y

#pH
p9_a <- ggplot(aes(x = pH, fill= quality.cat), data = redwines) + geom_histogram(binwidth=.01, color = "black", size= 0.25) +  scale_x_log10() + ggtitle('pH') + theme_light(base_size = 10)
p9_b <-ggplot(aes(y = pH, x= quality.cat, fill= quality.cat), data = redwines) + geom_boxplot(alpha= 1, outlier.alpha = 1/10) + scale_y_log10() + ggtitle('pH')+ theme_light(base_size = 10)#transformed y

#Sulphates
p10_a <- ggplot(aes(x = sulphates, fill= quality.cat), data = redwines) + geom_histogram(binwidth=.05, color = "black", size= 0.25) +  scale_x_log10() + ggtitle('Sulphates')+ theme_light(base_size = 10)
p10_b <-ggplot(aes(y = sulphates, x= quality.cat, fill= quality.cat), data = redwines) + geom_boxplot(alpha= 1, outlier.alpha = 1/10) + scale_y_log10() + ggtitle('Sulphates')+  theme_light(base_size = 10)#transformed y

#Alcohol
p11_a <- ggplot(aes(x = alcohol, fill= quality.cat), data = redwines) + geom_histogram(binwidth=.01, color = "black", size= 0.25) +  scale_x_log10() + ggtitle('Alcohol')+ theme_light(base_size = 10)
p11_b <-ggplot(aes(y = alcohol, x= quality.cat, fill= quality.cat), data = redwines) + geom_boxplot(alpha= 1, outlier.alpha = 1/10) + scale_y_log10() + ggtitle('Alcohol') + theme_light(base_size = 10)#transformed y

grid.arrange(p8_a,p8_b,p9_a,p9_b,p10_a,p10_b,p11_a,p11_b, ncol=2) + theme_light(base_size = 10);
```

# Bivariate Analysis

### Talk about some of the relationships you observed in this part of the investigation. How did the feature(s) of interest vary with other features in the dataset?

### Did you observe any interesting relationships between the other features (not the main feature(s) of interest)?

### What was the strongest relationship you found?




# Multivariate Plots Section

## Correlation betweeen variables


```{r echo=FALSE,include= TRUE, Multivariate_Plots_2, fig.width=5, fig.height=5, fig.align= 'center'}


ggcorr_custom <- function(x, method = "pairwise", palette = "RdYlGn", cols_to_avoid = c("X", "quality.cat"), layot_exp = 0) {
  require(ggplot2)
  require(GGally)
  require(grid)
  require(gridBase)
  require(gridExtra)

  column_names = colnames(x)
  cols_names = column_names[!(column_names %in% cols_to_avoid)]

  po.nopanel <- list(theme(
    panel.background = element_blank(), 
    panel.grid.minor = element_blank(),
    panel.grid.major = element_blank(),
    axis.text.x = element_text(angle = -90)))
  
  corr <-ggcorr(data=subset(x, select = cols_names),
         label = TRUE,
         geom = "blank",
         hjust = 1,
         angle = -25,
         digits = 2,
         nbreaks = 8,
         name = "",
         label_round = 1,
         label_size = 3,
         legend.size = 7,
         layout.exp = layot_exp,
         size = 3 #argument of geom_text for the diagonals
         ) +
    geom_point(aes(size=abs(coefficient), color = breaks, 
                   alpha = round((abs(coefficient) > 0.28)*abs(coefficient), 1))) +
    scale_color_brewer(palette = "RdBu")+
    scale_alpha(range=c(0.,.7))+
    scale_size(range=c(0,9))+
    guides(fill= FALSE, color= FALSE, size = FALSE, alpha = FALSE)
return(corr)
}

ggcorr_custom(redwines, layot_exp = 2) + ggplot2::theme_light(base_size= 7)

```

## Correlation by quality
```{r, fig.width=12}

corr_total_1 <- ggcorr_custom(subset(redwines, quality.cat=="bad"), layot_exp = 2) + ggplot2::theme_light(base_size= 8) + ggplot2:: ggtitle("Bad")

corr_total_2 <- ggcorr_custom(subset(redwines, quality.cat=="medium"), layot_exp = 2) + ggplot2::theme_light(base_size= 8) + ggplot2:: ggtitle("Medium")

corr_total_3 <- ggcorr_custom(subset(redwines, quality.cat=="good"), layot_exp = 2) + ggplot2::theme_light(base_size= 8)  + ggplot2:: ggtitle("Good")



gridExtra::grid.arrange(corr_total_1, corr_total_2, corr_total_3, ncol = 3, name = "Correlation vs Quality") + ggplot2:: theme_light(base_size = 6)
#temp <- gridExtra::arrangeGrob(corr_total_1, corr_total_2, corr_total_3, nrow = 3, name = "Correlation vs Quality") + ggplot2:: theme_light(base_size = 10)

#grid.newpage()
#grid.draw(temp)

```

From the general correlation matrix it three main variables can be selected due to their high correlation with quality. In order to see if they are suitable to perform an analysis let us explore the correlation between them.

```{r Correlations_comparative_from_table, fig.width=3, fig.align= 'center'}


ggcorr_custom(subset(redwines, select= c("volatile.acidity", "alcohol", "sulphates")), layot_exp = 1 ) + ggplot2::theme_light(base_size= 8)
#c("volatile.acidity", "alcohol", "sulphates", "quality.cat","quality"))

multi1 <- ggplot(aes(x = sulphates, y = alcohol, fill= quality.cat), data = redwines)+ geom_point(alpha=1/5, aes(color= quality.cat))+ facet_wrap(~redwines$quality.cat) +theme_light(base_size = 8)



multi2 <- ggplot(aes(x = sulphates, y = volatile.acidity, fill= quality.cat), data = redwines)+ geom_point(alpha=1/5, aes(color= quality.cat))+ facet_wrap(~redwines$quality.cat) +theme_light(base_size = 8)



multi3 <- ggplot(aes(x = volatile.acidity, y = alcohol, fill= quality.cat), data = redwines)+ geom_point(alpha=1/5, aes(color= quality.cat))+ facet_wrap(~redwines$quality.cat) +theme_light(base_size = 8)


grid.arrange(multi1,multi2,multi3, ncol=1) + theme_light(base_size = 10)#+ ggplot2::guides(color= FALSE, fill=FALSE);

```


```{r Correlations_comparative_from_article, fig.width=3, fig.align= 'center'}

ggcorr_custom(subset(redwines, select= c("total.sulfur.dioxide","pH", "sulphates")) , layot_exp = 1) + ggplot2::theme_light(base_size= 8)
#c("total.sulfur.dioxide","pH", "sulphates", "quality.cat","quality")

multi1 <- ggplot(aes(x = pH, y = sulphates, fill= quality.cat), data = redwines)+ geom_point(alpha=1/5, aes(color= quality.cat))+ facet_wrap(~redwines$quality.cat) +theme_light(base_size = 8)



multi2 <- ggplot(aes(x = pH, y = total.sulfur.dioxide, fill= quality.cat), data = redwines)+ geom_point(alpha=1/5, aes(color= quality.cat))+ facet_wrap(~redwines$quality.cat) +theme_light(base_size = 8)



multi3 <- ggplot(aes(x = sulphates, y = fixed.acidity, fill= quality.cat), data = redwines)+ geom_point(alpha=1/5, aes(color= quality.cat))+ facet_wrap(~redwines$quality.cat) +theme_light(base_size = 8)


grid.newpage()
grid.arrange(multi1,multi2,multi3, ncol=1) + theme_light(base_size = 10);

```






##########################
```{r echo=FALSE, include=FALSE, Multivariate_Plots}
require(GGally)
require(data.table)
tmp <- data.table(subset(redwines, quality.cat=="good"), keep.rownames=TRUE, keep.colnames= TRUE)

d <- ggpairs(data=tmp, 
             
             #diag=list(continuous="density"),
             lower= list(continuous = wrap("points", alpha = 1/5)),
             columns=c("sulphates","pH","total.sulfur.dioxide"),
             mapping = ggplot2::aes(color = quality.cat), 
             axisLabels="show") + 
    theme_light(base_size = 10)
d
#d + facet_grid(~quality.cat)
```


# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the investigation. Were there features that strengthened each other in terms of looking at your feature(s) of interest?

### Were there any interesting or surprising interactions between features?

### OPTIONAL: Did you create any models with your dataset? Discuss the strengths and limitations of your model.

------

# Final Plots and Summary

### Plot One
```{r echo=FALSE, Plot_One}

```

### Description One


### Plot Two
```{r echo=FALSE, Plot_Two}

```

### Description Two


### Plot Three
```{r echo=FALSE, Plot_Three}

```

### Description Three

------

# Reflection


## References









```{r echo=FALSE, include= FALSE, Univariate_Plots_log}

p1 <-ggplot(aes(x = fixed.acidity ), data = redwines) + geom_histogram(binwidth = .01) +  scale_x_log10() #+ theme_light(base_size = 10)

p2 <-ggplot(aes(x = volatile.acidity ), data = redwines) + geom_histogram(binwidth = .01)+  scale_x_log10() #+ theme_light(base_size = 10)

p3 <-ggplot(aes(x = citric.acid ), data = redwines) + geom_histogram(binwidth = .01)#+ theme_light(base_size = 10)

p4 <-ggplot(aes(x = residual.sugar), data = redwines) + geom_histogram(binwidth = .05) +  scale_x_log10() #+ theme_light(base_size = 10)

p5 <-ggplot(aes(x = chlorides ), data = redwines) + geom_histogram(binwidth = .01)  +  scale_x_log10()#+ theme_light

p6 <-ggplot(aes(x = free.sulfur.dioxide), data = redwines) + geom_histogram(binwidth = 0.05) +  scale_x_log10() #+ theme_light(base_size = 10)

p7 <-ggplot(aes(x = total.sulfur.dioxide), data = redwines) + geom_histogram(binwidth = 0.05) #+ theme_light(base_size = 10)

p8 <-ggplot(aes(x = density), data = redwines) + geom_histogram(binwidth = .0005) #+ theme_light(base_size = 10)



grid.arrange(p1,p2,p3,p4,p5,p6,p7,p8, ncol=4) + theme_light(base_size = 10)
```

```{r echo=FALSE, include= TRUE, message= FALSE, warning= FALSE, Univariate_Plots_1, fig.width=16, fig.height=8}
library(gridExtra)
p1 <-ggplot(aes(x = fixed.acidity ), data = redwines) + geom_histogram(binwidth=.5)# +  scale_x_log10() #+ theme_light(base_size = 10)

p2 <-ggplot(aes(x = volatile.acidity ), data = redwines) + geom_histogram(binwidth = .05)#+  scale_x_log10() #+ theme_light(base_size = 10)

p3 <-ggplot(aes(x = citric.acid ), data = redwines) + geom_histogram(binwidth = .05)#+ theme_light(base_size = 10)

p4 <-ggplot(aes(x = residual.sugar), data = redwines) + geom_histogram(binwidth = .5) #+  scale_x_log10() #+ theme_light(base_size = 10)

p5 <-ggplot(aes(x = chlorides ), data = redwines) + geom_histogram(binwidth = .01)  #+  scale_x_log10()#+ theme_light

p6 <-ggplot(aes(x = free.sulfur.dioxide), data = redwines) + geom_histogram(binwidth = 5) #+  scale_x_log10() #+ theme_light(base_size = 10)

p7 <-ggplot(aes(x = total.sulfur.dioxide), data = redwines) + geom_histogram(binwidth = 10) #+ theme_light(base_size = 10)

p8 <-ggplot(aes(x = density), data = redwines) + geom_histogram(binwidth = .0005) #+ theme_light(base_size = 10)

p9 <-ggplot(aes(x = pH ), data = redwines) + geom_histogram(binwidth = .05)  #+  scale_x_log10()#+ theme_light

p10 <-ggplot(aes(x = sulphates), data = redwines) + geom_histogram(binwidth = 0.05) #+  scale_x_log10() #+ theme_light(base_size = 10)

p11 <-ggplot(aes(x = alcohol), data = redwines) + geom_histogram(binwidth = .5) #+ theme_light(base_size = 10)

p12 <-ggplot(aes(x = quality), data = redwines) + geom_bar() #+ theme_light(base_size = 10)

grid.arrange(p1,p2,p3,p4,p5,p6,p7,p8,p9,p10,p11,p12, ncol=4) + theme_light(base_size = 10)
```
