---
title: 'Project 4  : Exploration of Red Wine Quality'
author: "Isabel María Villalba Jiménez"
date: "December 6th, 2016"
output:
  bookdown::html_document2: 
    fig_caption: yes
  github_document:
    depth: 1
    fig_caption: yes
    fig_height: 6
    fig_width: 8
  html_document:
    depth: 1
    fig_caption: yes
    fig_height: 6
    fig_width: 8
  pdf_document:
    citation_package: natbib
    fig_caption: yes
    fig_height: 6
    fig_width: 7
    highlight: tango
    keep_tex: yes
link-citations: yes
bibliography: P4.bib
biblio-style: apsr
nocite: | 
  @jasonmedina, @SanjayMenon, @IwanThomas
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.cap = NULL, fig.width=4, fig.height=4, 
                      fig.path='Figs/', fig.align = 'center',
                      echo=FALSE, warning=FALSE, message=FALSE,
                      include = TRUE, results= 'hide', 
                      tidy.opts=list(width.cutoff=80),tidy=TRUE)
# Set so that long lines in R will be wrapped:
```

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
# Load all of the packages that you end up using
# in your analysis in this code chunk.


# Notice that the parameter "echo" was set to FALSE for this code chunk.
# This prevents the code from displaying in the knitted HTML output.
# You should set echo=FALSE for all code chunks in your file.
this_dir <- 'Git/data-analyst-nanodegree/P4-explore-and-summarize-data/' 
setwd("~")
setwd(this_dir)

#remove.packages("UpSetR")
#remove.packages("GGally")
require(ggplot2)
require(GGally)
library(gridExtra)
# stable version on CRAN
#install.packages("bookdown")
library(bookdown)



# Captioner
# Now included in bookdown
#captioner is an R package for generating figure numbers and captions
#https://github.com/adletaw/captioner

#install.packages("devtools")
#devtools::install_github("adletaw/captioner")
#library(captioner)

#Set default theme
theme_set(theme_light())
```


#Analysis

In this work it will be analyzed the impact in quality of several parameters describing red wine. The dataset is curated by Udacity and comes from UCI repository [https://archive.ics.uci.edu/ml/datasets/Wine+Quality](https://archive.ics.uci.edu/ml/datasets/Wine+Quality) and consists of 1599 sample data for Red wine [https://docs.google.com/document/d/1qEcwltBMlRYZT-l699-71TzInWfk4W9q5rTCSvDVMpc/pub?embedded=true](https://docs.google.com/document/d/1qEcwltBMlRYZT-l699-71TzInWfk4W9q5rTCSvDVMpc/pub?embedded=true).

 In @cortez2009modeling it is shown that the most imporant features for assessing Red Wine quality are:
 
- sulphates

- pH

- total sulfur dioxide

# Variable summary


```{r Load_the_Data}
# Load the Data
redwines <- read.csv('wineQualityReds.csv')

#dim(redwines)
#names(redwines)
summary(redwines)

# New variables 
#redwines$quality.factor <- factor(redwines$quality)

redwines$quality.cat <- NA
redwines$quality.cat <- ifelse(redwines$quality>=7, 'good','medium')
redwines$quality.cat <- ifelse(redwines$quality<=4, 'bad',redwines$quality.cat) 
# if not, leave the previous values
redwines$quality.cat <- factor(redwines$quality.cat,
                               levels = list('bad', 'medium','good')) # set the order!

print("Variables after dividing into quality groups")
str(redwines) #summary of values for each variable

#unique(redwines$quality.cat)
```

The dataset has `r dim(redwines)[1]` wine entries and `r dim(redwines)[0]` features.

The summary of the dataset is the following 
```{r results='hold'} 
summary(redwines)
```

- `X`: is the row id
- `quality`: range is between 3 and 8
- `quality.cat`: generated to group quality in ranges: 0-4 -> BAD, 
4-7 -> MEDIUM, 7-10 -> GOOD. More information on the section "Analysis".
- All variables except total.sulfar.dioxide and free.sulfar.dioxide (integers), 
are continuous.
- total.sulfer.dioxide is the sum of free.sulfur.dioxide and bound forms: hence 
the two sulfur variables are related.
- volitale.acidity is acetic acid, different from tartaric or fixed.acidity and 
citric.acid. Acetic acid gives wine vinegar taste, while fixed acids do not 
easily evaporate. Citric acid is added to some wines for freshness or to increase
acidity. @jasonmedina

# Univariate analysis

In this section it will be analyzed each of the variables describing the wines.

## Plots Section
###Quality

The distribution of wine shows that most of wines have a quality between 5-6 
points.

```{r echo=FALSE, UnivariatePlots,fig.width=3, fig.height=3, fig.cap = "Quality distribution"}
ggplot(aes(x = quality), data = redwines) + 
geom_histogram(binwidth = 1, 
                fill = "brown", color= "black") +
ggtitle('Quality') + 
theme_light(base_size = 10)
 
#ggplot(aes(x = sulphates), data = redwines) + geom_histogram(binwidth = .02) 
#+ scale_x_log10() + ggtitle('Sulphates')

library(GGally)
#ggpairs(redwines, axisLabels = 'internal', lower = list(continous = 'points', 
#mapping = aes(shape= quality.cat, color=quality.cat), size = 8), #
#upper = list(continous = 'smooth', discrete = 'facetwrap', size = 8), 
#columns=2:10) + theme_light(base_size = 8)
```
